---
title: "computation"
format: html
editor: visual
---

## 

```{r}
library(tidyverse)
library(palmerpenguins)
library(dplyr)
```

## Data Wrangling Refresher

1.  Starting from penguins, write a single piped sequence in which you:
    1.  ONLY include penguins at dream and biscoe islands
    2.  remove the year and sex variables
    3.  add a new column body_mass_kg with penguin mass converted from g to kg
    4.  rename the island variable to location

```{r}
penguins_sub <- penguins %>% filter(island== "Dream" | island == "Bisoce") %>%                 select(-c("year","sex")) %>% 
  mutate(body_mass_kg = body_mass_g/1000) %>% 
  rename(location = island)
```

2.  Starting from penguins, write a single piped seuqnce to:

-   limit to adlie penguins

-   remove obs where flipper lenghth is NA

-   group by sex

-   create summary table that contains mean, sd, sample size of flipper length for male and female adlie peguins

```{r}
penguins_sub2 <- penguins %>% filter(species == "Adelie") %>% 
  drop_na(flipper_length_mm) %>% 
  group_by(sex) %>% 
  summarize(mean = mean(flipper_length_mm),
            std_dev = sd(flipper_length_mm),
            sample_size = n())

penguins %>% count(species,island, sex)
```

```{r}
animals<- data.frame(
  stringsAsFactors = FALSE,
          location = c("lagoon", "bluff", "creek", "oaks", "bluff"),
           species = c("bobcat", "coyote", "fox", "squirrel", "bobcat"),
          maturity = c("adult", "juvenile", "adult", "juvenile", "adult")
)
```

```{r}
sites<- data.frame(
  stringsAsFactors = FALSE,
          location = c("beach", "lagoon", "bluff", "oaks"),
    full_site_name = c("Goleta Beach","UCSB Lagoon",
                       "Ellwood Mesa","Fremont Campground"),
      jurisdiction = c("SB City", "UCSB", "SB City", "USFS")
)
```

Full Join: Uses full_join()

```{r}
full_join_example <- full_join(animals,sites)
```

```{r}
left_join_example<- left_join(animals,sites)
```

```{r}
inner_join_example <- inner_join(animals,sites)
```
